---
title: "MCA Metadata Preprocessing"
author: "Mervin M Fansler"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document: 
    code_folding: show
    toc: true
    toc_float: true
---

# Purpose

We reprocess the Mouse Cell Atlas 1.1 metadata to a CSV format that is ready to 
use with a processing pipeline for annotating a mouse UTRome.

# Initialization

## Libraries
```{r libs, message=FALSE, warning=FALSE}
library(magrittr)
library(tidyverse)
library(readxl)
```

## Parameters
```{r set_params}
set.seed(20220711)
FILE_MCA_1_1="metadata/MCA1.1_cell_info.xlsx"
```

## Functions

```{r methods}
str_clean <- . %>%
    str_to_lower %>%
    str_replace_all(fixed("+"), "pos") %>%
    str_replace_all(fixed(" "), "-") %>%
    str_replace(fixed("("), "-") %>%
    str_replace(fixed(")"), "-") %>%
    str_remove("-$")
```

# Data
## Loading
```{r load_data, message=FALSE}
df_mca1 <- read_xlsx(FILE_MCA_1_1) %>% 
    rename(cell_id=cellnames) %>%
    mutate(cell_id=str_replace(cell_id, fixed("Male.fetal.Gonad_1"), "FetalMaleGonad_1"))
```

## Preprocessing
```{r prepare_data}
df_mca1_clean <- df_mca1 %>%
    mutate(sample_id=str_remove(str_remove(cell_id, '\\.[ACGT]{18}$'), fixed(".")),
           bx=str_extract(cell_id, "[ACGT]{18}$"),
           cell_id=str_c(sample_id, ".", bx),
           tissue=str_remove(tissue, fixed(".")),
           celltype=str_clean(celltype),
           celltype_id=str_c(cluster, "-", celltype)) %>%
    select(cell_id, sample_id, bx, stage, tissue, cluster, celltype, celltype_id)
```

# Data Summary
## Cells by stage
```{r stage_summary, fig.width=3, fig.height=4}
df_mca1_clean %>%
    mutate(stage=factor(stage, levels=c("Fetal", "Neonatal", "Adult"))) %>%
    ggplot(aes(x=stage, fill=stage)) +
    geom_bar() +
    scale_fill_grey() +
    scale_y_continuous(expand=c(0,0,0.05,0)) +
    labs(x="Stage", y="Cells") +
    guides(fill='none') +
    theme_light()
```

## Cells by sample
```{r batch_summary, fig.width=4, fig.height=10}
df_mca1_clean %>%
    mutate(sample_id=fct_infreq(factor(sample_id))) %>%
    ggplot(aes(x=sample_id, fill=tissue)) +
    geom_bar() +
    scale_fill_grey() +
    scale_y_continuous(expand=c(0,0,0.05,0)) +
    coord_flip() +
    labs(x="Sample", y="Cells") +
    guides(fill='none') +
    theme_light()
```

## Cells by tissue
```{r tissue_summary, fig.width=4, fig.height=6}
df_mca1_clean %>%
    mutate(tissue=fct_infreq(factor(tissue))) %>%
    ggplot(aes(x=tissue)) +
    geom_bar(fill='grey') +
    scale_y_continuous(expand=c(0,0,0.05,0)) +
    coord_flip() +
    labs(x="Tissue", y="Cells") +
    theme_light()
```

## Cells by cell type
```{r celltype_summary, fig.width=4, fig.height=6}
df_mca1_clean %>%
    mutate(celltype=fct_infreq(factor(celltype))) %>%
    ggplot(aes(x=celltype)) +
    geom_bar(fill='grey') +
    scale_y_continuous(expand=c(0,0,0.05,0)) +
    coord_flip() +
    labs(x="Cell Type", y="Cells") +
    theme_light()
```

## Cells by cell type ID
```{r celltype_id_summary, fig.width=12, fig.height=4}
df_mca1_clean %>%
    mutate(celltype_id=fct_infreq(factor(celltype_id))) %>%
    ggplot(aes(x=celltype_id)) +
    geom_bar(fill='grey') +
    scale_x_discrete(guide=guide_axis(angle=90)) +
    scale_y_continuous(expand=c(0,0,0.05,0)) +
    labs(x="Cell Type ID", y="Cells") +
    theme_light()
```

# Export
```{r export}
df_mca1_clean %>% write_csv("metadata/mca1_1.cell_annot.csv.gz")

df_mca1_clean %>%
    select(celltype_id, sample_id) %>%
    distinct() %>%
    arrange(celltype_id) %>%
    write_csv("metadata/mca1_1.celltype_sample.csv")
```



---

# Runtime Details
## Session Info
```{r sesh_info, echo=FALSE}
sessionInfo()
```

## Conda Environment
```{bash comment="", echo=FALSE}
if ! command -v conda &> /dev/null
then
  echo "Conda not detected."
elif [ -z "${CONDA_PREFIX}" ]
then
  echo "No active Conda environment."
else
  echo "## Conda Environment YAML"
  conda env export
fi
```
