---
title: "Merged Length Distributions"
author: "Mervin M Fansler"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document: 
    code_folding: show
    toc: true
    toc_float: true
---

# Purpose

Compare the distributions of merged detection regions in the UTRome annotation 
across various parameters.

# Initialization

## Libraries
```{r libs, message=FALSE, warning=FALSE}
library(magrittr)
library(tidyverse)
library(cowplot)
```

## Parameters
```{r set_params}
set.seed(20220404)

FILE_PATH="qc/gff"
FILE_PATTERN="utrome.merged_lengths.*.tsv.gz"
```

## Functions

```{r methods}

```

# Data
## Loading
```{r load_data, message=FALSE}
df_lengths <- list.files(FILE_PATH, FILE_PATTERN, full.names=TRUE) %>%
    map(read_tsv) %>%
    do.call(what=rbind)
```

## Preprocessing
```{r prepare_data}

```

# Analysis

## Merge Count Distribution

```{r fig_merge_ct, fig.width=6, fig.height=5}
MAX_COUNT=6
MAX_LABEL="6+"
df_lengths %>%
    mutate(n_merged_fct=ifelse(n_merged >= MAX_COUNT, MAX_LABEL, as.character(n_merged))) %>%
    ggplot(aes(x=n_merged_fct)) +
    geom_bar(fill='grey', color='black') +
    facet_grid(rows=vars(min_tpm), cols=vars(epsilon)) +
    scale_y_continuous(expand=c(0,0,0,1000)) +
    labs(x="Overlapping Cleavage Sites Merged", y="Count", 
         title="Merging Distance vs Minimum TPM") +
    theme_bw()
```

## Merged Length Distribution
### Histogram - Raw
```{r hist_merge_len_raw, fig.width=6, fig.height=5}
df_lengths %>%
    ggplot(aes(x=nts_measured)) +
    geom_histogram(fill='grey', color='black', binwidth=50) +
    facet_grid(rows=vars(min_tpm), cols=vars(epsilon)) +
    scale_x_continuous(guide = guide_axis(angle = 90)) +
    scale_y_continuous(expand=c(0,0,0,1000)) +
    labs(x="Merged Length", y="Count", 
         title="Merging Distance vs Minimum TPM") +
    theme_bw()
```

### Histogram - Log10
```{r hist_merge_len_log, fig.width=6, fig.height=5}
df_lengths %>%
    ggplot(aes(x=nts_measured)) +
    geom_histogram(fill='grey', color='black', binwidth=50) +
    facet_grid(rows=vars(min_tpm), cols=vars(epsilon)) +
    scale_x_continuous(guide = guide_axis(angle = 90)) +
    scale_y_log10(expand=c(0,0,0,0.4)) +
    labs(x="Merged Length", y="Count", 
         title="Merging Distance vs Minimum TPM") +
    theme_bw()
```


### Histogram - Truncated
```{r hist_merge_len_trunc, fig.width=6, fig.height=5}
NT_BREAKS=c(-Inf, 500, 550,600,650,700,750,800,Inf)
NT_LABELS=c("< 500", "500-549", "550-599", "600-649", 
            "650-699", "700-749", "750-799", ">= 800")
df_lengths %>%
    ggplot(aes(x=cut(nts_measured, right=FALSE,
                     breaks=NT_BREAKS,
                     labels=NT_LABELS))) +
    geom_bar(fill='grey', color='black') +
    facet_grid(rows=vars(min_tpm), cols=vars(epsilon)) +
    scale_x_discrete(guide = guide_axis(angle = 90)) +
    scale_y_continuous(expand=c(0,0,0,1000)) +
    labs(x="Merged Length", y="Count", 
         title="Merging Distance vs Minimum TPM") +
    theme_bw()
```


### Histogram - Truncated Log10
```{r hist_merge_len_trunc_log, fig.width=6, fig.height=5}
NT_BREAKS=c(-Inf, 500, 550,600,650,700,750,800,Inf)
NT_LABELS=c("< 500", "500-549", "550-599", "600-649", 
            "650-699", "700-749", "750-799", ">= 800")
df_lengths %>%
    ggplot(aes(x=cut(nts_measured, right=FALSE,
                     breaks=NT_BREAKS,
                     labels=NT_LABELS))) +
    geom_bar(fill='grey', color='black') +
    facet_grid(rows=vars(min_tpm), cols=vars(epsilon)) +
    scale_x_discrete(guide = guide_axis(angle = 90)) +
    scale_y_log10(expand=c(0,0,0,0.4)) +
    labs(x="Merged Length", y="Count", 
         title="Merging Distance vs Minimum TPM") +
    theme_bw()
```


---

# Runtime Details
## Session Info
```{r sesh_info, echo=FALSE}
sessionInfo()
```

## Conda Environment
```{bash comment="", echo=FALSE}
if ! command -v conda &> /dev/null
then
  echo "Conda not detected."
elif [ -z "${CONDA_PREFIX}" ]
then
  echo "No active Conda environment."
else
  echo "## Conda Environment YAML"
  conda env export
fi
```
